---
title: "R Notebook to summarize the solutions for the extra exercises of R-Tutor"
output: html_notebook
---

You can find the original tutorials here: 
http://www.r-tutor.com/elementary-statistics 


# Qualitative Data

## Frequency Distribution of Qualitative Data 
http://www.r-tutor.com/elementary-statistics/qualitative-data/frequency-distribution-qualitative-data

```{r}
library(MASS)
head(painters)

```


1) Find the frequency distribution of the composition scores in painters. 

```{r}
composition = painters$Composition
composition.freq = table(composition)
cbind(composition.freq)

```


2) Find programmatically the school that has the most painters. 

```{r}
school = painters$School
school.freq = table(school)
most_painters = max(school.freq)
most_painters

```


## Relative Frequency Distribution of Qualitative Data
http://www.r-tutor.com/elementary-statistics/qualitative-data/relative-frequency-distribution-qualitative-data


1) Find the relative frequency distribution of the composition scores in painters.

```{r}
old=options(digits = 1) #setting significant digits to 1

composition.relfreq = composition.freq / nrow(painters)
cbind(composition.relfreq)

options(old) #resetting old settings
```

## Bar Graph
http://www.r-tutor.com/elementary-statistics/qualitative-data/bar-graph

```{r setting colors}
colors = c("red", "yellow", "green", "violet", "orange", "blue", "pink", "cyan") 
```


1) Find the bar graph of the composition scores in painters.

```{r}
barplot(composition.freq, col = colors)
```

## Pie Chart
http://www.r-tutor.com/elementary-statistics/qualitative-data/pie-chart

1) Find the pie chart of the composition scores in painters.

```{r}
pie(composition.freq, col = colors)
```

## Category Statistics
http://www.r-tutor.com/elementary-statistics/qualitative-data/category-statistics

```{r}
school = painters$School
c_school = school == "C" 
c_painters = painters[c_school, ]  # child data set
mean(c_painters$Composition) 
```
```{r}
tapply(painters$Composition, painters$School, mean) 
```


1) Find programmatically the school with the highest composition scores.

```{r}
max_comp = tapply(painters$Composition, painters$School, max)
max_comp[which(max_comp %in% c(max(max_comp)))]
```


2) Find the percentage of painters whose color score is equal to or above 14.

```{r}
above14 = sum(painters$Colour >= 14)
all_painters = nrow(painters)

perc = above14 / all_painters * 100
sprintf("Painters with a Colour Score above 14: %.2f %%", perc)
```

# Quantitative Data

```{r}
head(faithful)
```

## Frequency Distribution of Quantitative Data

1) Find the frequency distribution of the eruption waiting periods in faithful.

```{r}
waits = faithful$waiting
# range(waits) # min = 43; max = 96
wbreaks = seq(40, 100, by = 5)
waits.cut = cut(waits, wbreaks, right = FALSE)
waits.freq = table(waits.cut)
cbind(waits.freq)
```

2) Find programmatically the duration sub-interval that has the most eruptions.

```{r}
head(duration.freq[order(-duration.freq)],1)
```


## Histogram
```{r}
colors = c("red", "yellow", "green", "violet", "orange", "blue", "pink", "cyan") 
```

1) Find the histogram of the eruption waiting period in faithful.


```{r}
hist(waits, right = FALSE, col = colors, main = "Old Faithful Waiting durations", xlab = "Waits in minutes")
```

## Relative Frequency Distribution of Quantitative Data

1) Find the relative frequency distribution of the eruption waiting periods in faithful.

```{r}
old = options(digits = 1)

waits.relfreq = waits.freq / nrow(faithful)
#waits.relfreq
cbind(waits.freq, waits.relfreq)

options(old)
```

## Cumulative Frequency Distribution
https://www.r-tutor.com/elementary-statistics/quantitative-data/cumulative-frequency-distribution

1) Find the cumulative frequency distribution of the eruption waiting periods in faithful.

```{r}
waits.cumfreq = cumsum(waits.freq)
cbind(waits.freq, waits.cumfreq)
```

## Cumulative Frequency Graph

1) Find the cumulative frequency graph of the eruption waiting periods in faithful.


```{r}
cumfreq0_w = c(0, waits.cumfreq)
plot(wbreaks, cumfreq0_w, main = "Old Faithful Waiting times", xlab = "Waiting Minutes", ylab = "Cumulative Waiting")
lines(wbreaks,cumfreq0_w)
```

## Cumulative Relative Frequency Distribution

1) Find the cumulative frequency distribution of the eruption waiting periods in faithful.

```{r}
old = options(digits = 2)

waits.cumrelfreq = waits.cumfreq / nrow(faithful)
cbind(waits.cumfreq, waits.cumrelfreq)

options(old)
```

## Cumulative Relative Frequency Graph

1) Find the cumulative relative frequency graph of the eruption waiting periods in faithful.

```{r}
cumrelfreq0_w = c(0, waits.cumrelfreq)
plot(wbreaks, cumrelfreq0_w, main = "Old Faithful Waiting times", xlab = "Waiting minutes", ylab = "Cumulative waiting proportions")
lines(wbreaks, cumrelfreq0_w)
```

Alternatively: 
```{r}
Fn_w = ecdf(waits)
plot(Fn_w, main = "Old Faithful Waiting times", xlab = "Waiting minutes", ylab = "Cumulative waiting proportions")
```

## Stem-and-Leaf Plot

1) Find the stem-and-leaf plot of the eruption waiting periods in faithful.

```{r}
stem(waits)
```

## Scatter Plot

```{r}
plot(duration, waits, xlab = "Eruption duration", ylab = "Time waited", main = "Scatter plot of Durations vs Waiting times")
abline(lm(waits ~ duration))
```

# Numerical Measures
http://www.r-tutor.com/elementary-statistics/numerical-measures

1) Find the mean eruption waiting periods in faithful.

```{r}
mean(waits)
```

## Median

```{r}
median(waits)
```

## Quartile

1) Find the quartiles of the eruption waiting periods in faithful.

```{r}
quantile(waits)
```

## Percentile 

```{r}
quantile(waits, c(.17, .43, .67, .85))
```

## Range

1) Find the range of the eruption waiting periods in faithful.

```{r}
max(waits) - min(waits)
```

## Interquartile Range

1) Find the interquartile range of eruption waiting periods in faithful.


```{r}
IQR(waits)
```

## Box Plot

1) Find the box plot of the eruption waiting periods in faithful.

```{r}
boxplot(waits, horizontal = TRUE)
```

## Variance 

1) Find the variance of the eruption waiting periods in faithful.

```{r}
var(waits)
```

## Standard Deviation

1) Find the standard deviation of the eruption waiting periods in faithful.

```{r}
sd(waits)
```

## Covariance 

```{r}
cov(duration, waits)
```

## Correlation Coefficient

```{r}
cor(duration, waits)
```

## Central Moment

```{r}
# install.package("moments")
```


1) Find the third central moment of eruption waiting period in faithful.

```{r}
library(moments)
moment(waits, order = 3, central = TRUE)
```

## Skewness

1) Find the skewness of eruption waiting period in faithful.

```{r}
skewness(waits)
```

## Kurtosis

1) Find the excess kurtosis of eruption waiting period in faithful.


```{r}
kurtosis(waits)
```

# Probability Distributions
https://www.r-tutor.com/elementary-statistics/probability-distributions

```{r}
i=0
for (x in seq(0, 4)){i= i+ dbinom(x, size = 12, prob = 0.2)}
i

pbinom(4, size = 12, prob = 0.2)
ppois(16, lambda = 12)
ppois(16, lambda = 12, lower.tail = FALSE)

runif(10, min=1, max=3)
pexp(2, rate = 1/3)
pnorm(84, mean = 72, sd = 15.2, lower.tail = FALSE)
qchisq(.95, df = 7)
qt(c(.025, .975), df = 5)
qf(.95, df1 = 5, df2 = 2)
```


# Interval Estimation 
http://www.r-tutor.com/elementary-statistics/interval-estimation

```{r}
library(MASS)
```

