---
title: "R Notebook to summarize the solutions for the extra exercises of R-Tutor"
output: html_notebook
---

You can find the original tutorials here: 
http://www.r-tutor.com/elementary-statistics 


## Frequency Distribution of Qualitative Data 
http://www.r-tutor.com/elementary-statistics/qualitative-data/frequency-distribution-qualitative-data

```{r}
library(MASS)
head(painters)

```


1) Find the frequency distribution of the composition scores in painters. 

```{r}
composition = painters$Composition
composition.freq = table(composition)
cbind(composition.freq)

```


2) Find programmatically the school that has the most painters. 

```{r}
school = painters$School
school.freq = table(school)
most_painters = max(school.freq)
most_painters

```


## Relative Frequency Distribution of Qualitative Data
http://www.r-tutor.com/elementary-statistics/qualitative-data/relative-frequency-distribution-qualitative-data


1) Find the relative frequency distribution of the composition scores in painters.

```{r}
old=options(digits = 1) #setting significant digits to 1

composition.relfreq = composition.freq / nrow(painters)
cbind(composition.relfreq)

options(old) #resetting old settings
```

## Bar Graph
http://www.r-tutor.com/elementary-statistics/qualitative-data/bar-graph

```{r setting colors}
colors = c("red", "yellow", "green", "violet", "orange", "blue", "pink", "cyan") 
```


1) Find the bar graph of the composition scores in painters.

```{r}
barplot(composition.freq, col = colors)
```

## Pie Chart
http://www.r-tutor.com/elementary-statistics/qualitative-data/pie-chart

1) Find the pie chart of the composition scores in painters.

```{r}
pie(composition.freq, col = colors)
```

## Category Statistics
http://www.r-tutor.com/elementary-statistics/qualitative-data/category-statistics

```{r}
school = painters$School
c_school = school == "C" 
c_painters = painters[c_school, ]  # child data set
mean(c_painters$Composition) 
```
```{r}
tapply(painters$Composition, painters$School, mean) 
```


1) Find programmatically the school with the highest composition scores.

```{r}
max_comp = tapply(painters$Composition, painters$School, max)
max_comp[which(max_comp %in% c(max(max_comp)))]
```


2) Find the percentage of painters whose color score is equal to or above 14.

```{r}
above14 = sum(painters$Colour >= 14)
all_painters = nrow(painters)

perc = above14 / all_painters * 100
sprintf("Painters with a Colour Score above 14: %.2f %%", perc)
```

